---
title: "Shiny - FTI 2019"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---


## What is Shiny?

Shiny is an R package that makes it easy to build interactive web applications (apps) straight from R.

Let's look at some Shiny examples:
<br>
<a href="https://shiny.rstudio.com/gallery/telephones-by-region.html" target="_blank">Telephones by region</a>
<br>
<a href="https://shiny.rstudio.com/gallery/single-file-shiny-app.html" target="_blank">Random histogram</a>
<br>
<a href="https://shiny.rstudio.com/gallery/lego-set.html" target="_blank">Lego Set Database</a>



## Let's look under the hood of a Shiny App

Before doing anything, we need to load the Shiny library (click the play button to run the code):

```{r}
library(shiny)
```

A Shiny app has three main components:

* a user interface (ui) object

* a server function

* a call to the shinyApp function

Below is the code for a super basic Shiny app

```{r warning = FALSE}
ui <- basicPage("This is a real Shiny app") # the user interface

server <- function(input, output, session) { } #the server

shinyApp(ui = ui, server = server) # this launches your app
```

See if you can replicate this, except rename the variables ui to myUi, and server to myServer. Retype the code in the code block below, and run it (you don't need to type the comments)

```{r}

```


Ok, we have our first Shiny app! But it isn't very good. Let's expand on the user interface by creating some different areas:

```{r}
ui <- fluidPage(
  titlePanel("title panel"),

  sidebarLayout(
    sidebarPanel("sidebar panel"),
    mainPanel("main panel")
  )
)


server <- function(input, output, session) { } 

shinyApp(ui = ui, server = server) 

```

Try replicating this code in the box below, but change the title panel to say "Hello World!" instead of "title panel"

```{r}

```

## Styling the user interface (UI)

We can add content to our Shiny app by adding text to our panels, as you've just done by changing the contents of the title panel. In order to have more control over the appearance and functionality of our content, we can employ some basic HTML elements. 

Headings and paragraphs are two basic but very useful HTML tags.

Headings tags range from h1 to h6, h1 being the largest. Headings are bold by default, as opposed to paragraphs, which will just display regular text. The code below shows the syntax for adding headings and paragraphs to an app:

```{r}
ui <- fluidPage(

  titlePanel("title panel"),

  sidebarLayout(
    sidebarPanel("sidebar panel"),
    mainPanel(
      h1("This is an h1 heading"),
      h2("This is an h2 heading"),
      h3("This is an h3 heading"),
      h4("This is an h4 heading"),
      h5("This is an h5 heading"),
      h6("This is an h6 heading"),
        
      p("This is a paragraph")
    )
  )  
  

)

server <- function(input, output, session) { } 
shinyApp(ui = ui, server = server) 
```

Try creating an h1 tag below that says "Hello World!"

```{r}
ui <- fluidPage(
  titlePanel("title panel"),

  sidebarLayout(
    sidebarPanel("sidebar panel"),
    mainPanel(
      # add your code here
    )
  ) 
)

server <- function(input, output, session) { } 
shinyApp(ui = ui, server = server) 
```


We can also use an HTML tag to create links on our Shiny app. The code below shows the syntax for adding a link:

```{r}
ui <- fluidPage(

  titlePanel("title panel"),

  sidebarLayout(
    sidebarPanel("sidebar panel"),
    mainPanel(
      a("This is a link", href="https://shiny.rstudio.com/tutorial/written-tutorial/lesson1/")
    )
  )   
  
)

server <- function(input, output, session) { } 
shinyApp(ui = ui, server = server) 
```

Try changing the link below to redirect to https://www.lclark.edu

```{r}
ui <- fluidPage(

  titlePanel("title panel"),

  sidebarLayout(
    sidebarPanel("sidebar panel"),
    mainPanel(
      a("This is a link", href="https://shiny.rstudio.com/tutorial/written-tutorial/lesson1/")
    )
  )   
  
)

server <- function(input, output, session) { } 
shinyApp(ui = ui, server = server) 
```


## Adding user inputs to the UI

There are a number of widgets we can add to our Shiny app for users to interact with. For now, we'll discuss a couple of basic examples: the dropdown selector and the slider input.

**Example code for a dropdown selector:**
```{r}
ui <- fluidPage(

  titlePanel("title panel"),

  sidebarLayout(
    sidebarPanel(
      selectInput("select", h4("select box"), 
                choices = list("Choice 1" = 1, "Choice 2" = 2, "Choice 3" = 3), selected = 1) 
    
    # the 'selected' option defines which option is selected by default
    ),
    mainPanel(
      h1("my main panel")
    )
  ) 
  

)

server <- function(input, output, session) { } 
shinyApp(ui = ui, server = server) 
```

Now try altering the code below so the choices read "red", "blue", and "yellow", and option 3 is selected by default:
```{r}
ui <- fluidPage(
  titlePanel("title panel"),

  sidebarLayout(
    sidebarPanel(
      selectInput("select", h4("select box"), 
                choices = list("Choice 1" = 1, "Choice 2" = 2, "Choice 3" = 3), selected = 1) 
    
    ),
    mainPanel(
      h1("my main panel")
    )
  )
)

server <- function(input, output, session) { } 
shinyApp(ui = ui, server = server) 
```


**Example code for a slider input:**
```{r}
ui <- fluidPage(

  titlePanel("title panel"),

  sidebarLayout(
    sidebarPanel(
      sliderInput("slider", h4("slider"), min = 0, max = 100, value = 50)
    
      # the 'value' option defines the default value for the slider
    ),
    mainPanel(
      h1("my main panel")
    )
  )  
  
)

server <- function(input, output, session) { } 
shinyApp(ui = ui, server = server) 
```

Now try altering the code below so the slider ranges between 50 and 500, and 100 is the default value:
```{r}
ui <- fluidPage(

  titlePanel("title panel"),

  sidebarLayout(
    sidebarPanel(
      sliderInput("slider", h4("slider"), min = 0, max = 100, value = 50)
    
      # the 'value' option defines the default value for the slider
    ),
    mainPanel(
      h1("my main panel")
    )
  )  
  
)

server <- function(input, output, session) { } 
shinyApp(ui = ui, server = server) 
```



## Passing data from the user inputs to the main panel

In order to pass data from our user input to our display on the main panel, we need to do two things:
* add an R object to the ui
* write a short subroutine in the server function, which handles the inputs and outputs for our app

Here is an example of what that might look like:
```{r}
ui <- fluidPage(
  titlePanel("title panel"),
  sidebarLayout(
    sidebarPanel(
      # creating our dropdown and naming it 'dropdown1'
      selectInput("dropdown1", h4("select box"), 
                  choices = list(
                    "Choice 1" = 1, 
                    "Choice 2" = 2, 
                    "Choice 3" = 3,
                    "Choice 4" = 4,
                    "Choice 5" = 5
                    ), 
                  selected = 1) # option 1 is selected by default
    ),
    
    mainPanel(
      # calling a function which we've written in the server function
      h4( # function call is wrapped in h4 tags
        # here is the R object
        textOutput("dropdown_selected") 
        )
    )
  )
)

server <- function(input, output, session) {
    # function to display selected option
    output$dropdown_selected <- renderText({ 
      paste("You have selected", input$dropdown1) # displaying current value of dropdown1
  })
} 
shinyApp(ui = ui, server = server) 
```

Notice that the value you specify in the textOutput function ("dropdown_selected") matches the value after the output$ in the server function. You can name it whatever you want, as long as they match.

Similarly, notice that the same is true for the identifier of your ui selectInput ("dropdown1"). It matches the value after input$ in the server function.


Now try copying the code into the box below and changing the output message to "Great job, you've selected" followed by the current value of 'dropdown1'
```{r}

```
If you need something else to do, try changing the main panel output to display as h2 instead of h4


## Creating a basic histogram

Now we're going to branch out and plot a histogram with some user input. We'll use a sliderInput for our user input. Our R object in the mainPanel is "plotOutput", which is different from the "textOuput" function we used in the last example. Shiny provides a list of possible functions (here)[https://shiny.rstudio.com/tutorial/written-tutorial/lesson4/], depending upon what you want your output to be. Similarly, in our server function, we use "renderPlot" instead of "renderText". The previous link also provides a list of available functions for the server code. 



```{r}
ui <- fluidPage(
  titlePanel("Make a Histogram!"),
  
  sidebarLayout(
    sidebarPanel(
      helpText("Adjust the slider to update the histogram"),

      
      sliderInput("slider", 
                  label = "Pick a value:",
                  min = 10, max = 100, value = c(10))
    ),
    
    mainPanel(
      plotOutput("my_histogram")
    )
  )
)

server <- function(input, output) {

  output$my_histogram <-renderPlot({
    # first define our input value from the slider as myNumber
    myNumber<-input$slider[1]
    #run the hist function. We're also adding the input value as part of the x-label value, using the paste function
    hist(runif(myNumber), main="My Histogram", xlab=paste(myNumber, " total samples"))

  })
  
}
shinyApp(ui = ui, server = server) 

```



## Using a dataset

Now let's see if we can create a Shiny app that uses a dataset. We're going to use the publicly available dataset on alcohol consumption by country, from [fivethirtyeight's github repository](https://github.com/fivethirtyeight/data/tree/master/alcohol-consumption). This file is already loaded in our workspace, so run the following commands, and take a look at the dataframe:


```{r}
drinks=read.csv("drinks.csv")

drinks
#units are "servings per person in the year 2010"
```


Let's use this data set to create a Shiny app! This example is a little more advanced, so let's walk through it below. 


```{r}
#for data filtering later
library(tidyverse)

# drinks data frame
drinks=read.csv("drinks.csv")

# this creates the data for the pulldown menu
title <- as.vector(drinks$country)
code <- as.vector(drinks$country)
#maps code values to title values 
names(code) <- title


ui <- fluidPage(
  titlePanel("Alcohol Consumption by Country"),
  
  sidebarLayout(
    sidebarPanel(
      ("Data from the World Health Organization, 2010."),
      
      # in our select menu, we're using the code variable defined above
      selectInput("country", "Select a country",  choices=code)
    ),
    
    mainPanel(
      plotOutput("selected_country")
    )
  )
)

server <- function(input, output) {

  output$selected_country <- renderPlot({
    # first get the table row (myData) in which the country value in the data table matches our input (user selected) country
    myData<-drinks %>% filter(drinks$country==input$country)
    
    #get the specific values for each type of alcohol
    beer<-myData$beer_servings
    wine<-myData$wine_servings
    spirits<-myData$spirit_servings
    
    #set the x-axis labels
    labels <- c("beer", "wine", "sprits")
    
    #generate the barplot
    barplot(c(beer, wine, spirits),names.arg=labels, main=input$country,  ylab="servings/year", col="blue")

  })

}
shinyApp(ui = ui, server = server) 



```



Now copy the code above and paste it below. Try editing it to do the following:
* make a barplot with just a single bar: the total litres of pure alcohol 
* change the bar color
* use the paste() function to make a more descriptive plot title (hint: this is the "main" parameter)

```{r}

```



